---
title: "DAPC"
author: "Zhian N. Kamvar"
date: "September 29, 2014"
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
    toc_depth: 2
---

Purpose
-------

This document will explore the DAPC analyses for the data given different
populations.

Required packages and data
--------

```{r}
library(PramCurry)
library(reshape2)
library(ggplot2)
library(ape)
library(poppr)
library(adegenet)
library(igraph)
options(stringsAsFactors = FALSE)
data(ramdat)
data(pop_data)
data(myPal)
sessionInfo()
```

DAPC analysis on ZONE2
-----

```{r}
# plot_posterior <- function(da.object, gid, pal){
#   posterior <- da.object$posterior
#   names(dimnames(posterior)) <- c("sample", "population")
#   to_merge <- data.frame(list(sample = dimnames(posterior)$sample, 
#                               oldPopulation = pop(gid)))
#   post <- melt(posterior, value.name = "probability")
#   post <- merge(post, to_merge)
#   if (is.numeric(post$sample)){
#     post$sample <- factor(post$sample, levels = unique(post$sample))
#   }
#   outPlot <- ggplot(post, aes(x = sample, fill = population, y = probability)) + 
#     geom_bar(stat = "identity", position = "fill", width = 1) + 
#     theme_classic() + 
#     theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
#     scale_y_continuous(expand = c(0, 0)) +
#     scale_x_discrete(expand = c(0, 0)) +
#     facet_wrap(~oldPopulation, scales = "free_x", drop = TRUE, ncol = 1) +
#     scale_fill_manual(values = pal)
#   return(outPlot)
# }
```




I have originally performed these by utilizing the `adegenetServer()` function.
Since it is interactive, I will only post the code to do it here:

```{r, eval=FALSE}
setpop(ramdat) <- ~ZONE2
save(ramdat, file = "zone2.rda")
adegenetServer()
```

This will give me the tools to look at things like the composition plot and the
cross-validation plot.


### Cross-validation

```{r, Z2xval, cache = TRUE}
setpop(ramdat) <- ~ZONE2
xval <- xvalDapc(ramdat@tab, pop(ramdat), n.pca.max = 15, result = "overall", n.rep = 1000)
xval[-1]
```

### Running and plotting

```{r}
setpop(ramdat) <- ~ZONE2
(z2.dapc <- dapc(ramdat, n.pca = 12, n.da = 6))
```

```{r, fig.height = 10, fig.width = 10}
loadingplot(z2.dapc$var.contr, axis = 1)
loadingplot(z2.dapc$var.contr, axis = 2)
png("loading_figure.png", width = 183, height = 88, units = "mm", res = 300)
loadingplot(z2.dapc$var.contr, axis = 1)
dev.off()
popPal <- char2pal(ramdat@pop.names)
scatter(z2.dapc, cex = 2, legend = TRUE, clabel = FALSE, col = popPal,
        posi.leg = "bottomleft", scree.pca = TRUE, posi.pca = "topright",
        posi.da = "bottomright", cleg = 0.75, xax = 1, yax = 2, inset.solid = 1, 
        ratio.pca = 0.2, ratio.da = 0.2)
png("no_nursery_DAPC.png", width = 183, height = 183, units = "mm", res = 300)
scatter(z2.dapc, cex = 2, legend = TRUE, clabel = FALSE, col = popPal,
        posi.leg = "bottomright", scree.pca = TRUE, posi.pca = "topright",
        posi.da = "bottomleft", cleg = 0.75, xax = 1, yax = 2, inset.solid = 1, 
        ratio.pca = 0.2, ratio.da = 0.2)
dev.off()
```


```{r, fig.height = 20, fig.width = 13}
plot_posterior(z2.dapc, ramdat, popPal)
```



### Removing biases

Here, I am removing the Cape Sebastian isolates as well as the populations with
less than 10 isolates.

```{r, nosebXval, cache = TRUE}
noseb <- popsub(ramdat, blacklist = "HunterCr", drop = FALSE)
noseb.gt.10 <- selPopSize(noseb, n = 10)
# save(noseb.gt.10, file = "nosebpr.rda")
noseb.xval <- xvalDapc(noseb.gt.10@tab, pop(noseb.gt.10), n.pca.max = 15, 
                       result = "overall", n.rep = 1000)
noseb.xval[-1]
```
```{r}
(noseb.dapc <- dapc(noseb.gt.10, n.pca = 15, n.da = 4))
```

```{r, fig.height = 10, fig.width = 10}
scatter(noseb.dapc, cex = 2, legend = TRUE, clabel = FALSE, 
        col = popPal[noseb.gt.10@pop.names], scree.pca = TRUE,
        posi.leg = "bottomleft", posi.pca = "topleft",
        cleg = 0.75)
```

```{r, fig.height = 20, fig.width = 13}
plot_posterior(noseb.dapc, noseb.gt.10, popPal)
```

### Predicting sources. 

Since we now have a model from our well represented populations, we can use this 
to predict the sources of individuals from these populations. Let's see if they
match up


```{r, fig.height = 10, fig.width = 10}
to_test <- noseb@pop.names[!noseb@pop.names %in% noseb.gt.10@pop.names]
testpop <- popsub(noseb, to_test, drop = FALSE)
predictions <- predict.dapc(noseb.dapc, newdata = testpop)
plot_posterior(predictions, testpop, popPal)
```

```{r, fig.height = 10, fig.width = 10}
to_test <- ramdat@pop.names[!ramdat@pop.names %in% noseb.gt.10@pop.names]
testpop <- popsub(ramdat, to_test, drop = FALSE)
predictions <- predict.dapc(noseb.dapc, newdata = testpop)
plot_posterior(predictions, testpop, popPal)
```

Using 95% inertia ellipses.
```{r, fig.height = 10, fig.width = 10}
scatter(noseb.dapc, cex = 2, legend = TRUE, clabel = FALSE, cellipse = 2.5,
        col = popPal[noseb.gt.10@pop.names], scree.pca = FALSE,
        posi.leg = "bottomleft", posi.pca = "topleft",
        cleg = 0.75, bg="white", scree.da=0, pch=20, 
        xlim = c(-3, 7), ylim = c(-3, 3.5), solid = 0.5)
par(xpd=TRUE)
# points(noseb.dapc$ind.coord[, 1], noseb.dapc$ind.coord[, 2], pch=20,
#        col = transp(popPal[noseb.gt.10@pop.names]), cex = 2)
points(predictions$ind.scores[, 1], predictions$ind.scores[, 2], 
       pch = 20, col = transp("black", 0.2), cex = 3)
points(predictions$ind.scores[, 1], predictions$ind.scores[, 2], 
       pch = as.character(pop(testpop)),
       col = transp(popPal[as.character(pop(testpop))], 0.7), cex = 1)
add.scatter.eig(noseb.dapc$eig, 15, 1, 2, posi="bottomright", inset=.02)
sebpredmat <- matrix(c(
  rowMeans(apply(predictions$posterior[1:66, ], 1, function(x) x >= 0.5)),
  rowMeans(apply(predictions$posterior[1:66, ], 1, function(x) x >= 0.95)),
  rowMeans(apply(predictions$posterior[1:66, ], 1, function(x) x >= 0.99)),
  rowMeans(apply(predictions$posterior[1:66, ], 1, function(x) x >= 0.999))),
  ncol = 4)
dimnames(sebpredmat) <- list(Population = colnames(predictions$posterior),
                             `membership probability` = c("50%", "95%", "99%", 
                                                          "99.9%"))
signif(sebpredmat, 3)
```


