---
title: "Diversity_statistics"
author: "Zhian N. Kamvar"
date: "October 15, 2014"
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
    toc_depth: 2
---

Purpose
-------

This document will analyze and display the stastistics of diversity at the locus
and genotype level. 

Required packages and data
--------

```{r}
library(PramCurry)
library(reshape2)
library(ggplot2)
library(poppr)
library(adegenet)
library(mmod)
library(dplyr)
library(tidyr)
options(stringsAsFactors = FALSE)
data(ramdat)
data(pop_data)
data(myPal)
sessionInfo()
myTheme <- theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

Functions
------
```{r}
get_stats <- function(z){
  mat <- matrix(nrow = nrow(z), ncol = 4, 
                dimnames = list(Pop = rownames(z), 
                                Index = c("H", "G", "Hexp", "E.5")
                                )
                )
  N     <- rowSums(z)
  H     <- vegan::diversity(z)
  G     <- vegan::diversity(z, "inv")
  Simp  <- vegan::diversity(z, "simp")
  nei   <- (N/(N-1)) * Simp
  E.5   <- (G - 1)/(exp(H) -1)
  mat[] <- c(H, G, nei, E.5)
  return(mat)
}

boot_stats <- function(x, i){
  res        <- numeric(4)
  names(res) <- c("H", "G", "Hexp", "E.5")
  z     <- table(x[i])
  N     <- sum(z)
  H     <- vegan::diversity(z)
  G     <- vegan::diversity(z, "inv")
  Simp  <- vegan::diversity(z, "simp")
  nei   <- (N/(N-1)) * Simp
  E.5   <- (G - 1)/(exp(H) -1)
  res[] <- c(H, G, nei, E.5)
  return(res)
}

extract_samples <- function(x) rep(1:length(x), x)

do_boot <- function(tab, n, ...){
  res <- apply(tab, 1, function(x) boot::boot(extract_samples(x), boot_stats, n, ...))
  return(res)
}

get_ci <- function(x, lb, ub){
  res <- apply(x$t, 2, quantile, c(lb, ub), na.rm = TRUE)
  return(res)
}

get_all_ci <- function(res, ci = 95){
  lower_bound  <- (100 - ci)/200
  upper_bound  <- 1 - lower_bound
  funval       <- matrix(numeric(8), nrow = 2)
  CI           <- vapply(res, FUN = get_ci, FUN.VALUE = funval, 
                         lower_bound, upper_bound)
  dCI          <- dimnames(CI)
  dimnames(CI) <- list(CI    = dCI[[1]], 
                       Index = c("H", "G", "Hexp", "E.5"),
                       Pop   = dCI[[3]])
  
  return(CI)
}


boot_ci <- function(tab, n = 1000, ci = 95, total = TRUE, ...){
  if (!is.matrix(tab) & is.genind(tab)){
    tab <- mlg.table(tab, total = total, bar = FALSE)
  }
  res  <- do_boot(tab, n, ...)
  orig <- get_stats(tab)
  orig <- melt(orig)
  CI   <- get_all_ci(res, ci = ci)
  samp <- vapply(res, "[[", FUN.VALUE = res[[1]]$t, "t")
  dimnames(samp) <- list(NULL, 
                         Index = c("H", "G", "Hexp", "E.5"),
                         Pop = rownames(tab))
  pl <- ggplot(melt(samp), aes(x = factor(Pop), y = value, group = Pop)) + 
    geom_boxplot() + 
    geom_point(aes(color = factor(Pop), x = factor(Pop), y = value), 
               size = 5, pch = 16, data = orig) +
    xlab("Population") + labs(color = "Observed") +
    facet_wrap(~Index, scales = "free_y") + myTheme
  print(pl)
  return(CI)
}

calc_loc_table <- function(dat, hier, combine = FALSE, lev = "allele", cc = TRUE){
  if (cc){
    dat <- clonecorrect(dat, hier = hier, combine = combine)
  }
  theTable <- locus_table(dat, information = FALSE, lev = lev)
  theHier <- gethierarchy(dat, hier, combine = combine)
  if (combine){
    setpop(dat) <- hier
    theHier <- levels(theHier[[length(theHier)]])
  } else {
    setpop(dat) <- as.formula(paste0("~", names(theHier)[1]))
    theHier <- levels(theHier[[1]])
  }
  loc_tab <- lapply(theHier, 
                   function(x){
                     locus_table(dat, 
                                 population = x, 
                                 information = FALSE,
                                 lev = lev)
                   })
  names(loc_tab) <- theHier
  return(list(pops = loc_tab, total = theTable))
}


plot_loc_table <- function(tab){
#   tab <- theList[["pops"]]
#   tot <- melt(theList["total"])
  tab <- melt(tab)
  ggplot(tab, aes(y = value, x = L2, fill = L2, linetype = NA)) +
    geom_bar(stat = "identity") +
    geom_hline(aes(yintercept = value, linetype = L1), 
               data = tab[tab$L1 == "total", ]) + 
    facet_grid(summary ~ locus, scales = "free_y") +
    scale_linetype_discrete(labels = "", name = "Pooled") +
    myTheme
}



amova_pair <- function(pops, dat, hier, ...){
  dat <- popsub(dat, pops)
  if (any(table(pop(dat)) < 3)){
    return(NULL)
  }
  poppr.amova(dat, hier, ...)
}

pairwise_amova <- function(x, hier, ...){
  pops         <- x@pop.names
  pop_combs    <- combn(pops, 2)
  xlist        <- apply(pop_combs, 2, amova_pair, x, hier, ...)
  names(xlist) <- apply(pop_combs, 2, paste, collapse = " : ")
  return(xlist)
}

pairwise_amova_test <- function(pairlist, nrepet = 99){
  res <- lapply(names(pairlist), print_amova_pairs, pairlist, nrepet)
  names(res) <- names(pairlist)
  return(res)
}

print_amova_pairs <- function(pairname, pairlist, nrepet){
  thePair <- pairlist[[pairname]]
  if (!is.null(thePair)){
    cat(pairname, "\n")
    theTest <- randtest(thePair, nrepet = nrepet)
    try(plot(theTest))
    return(theTest)
  } else {
    return(NULL)
  }
}


getpval <- function(x){
  if (is.null(x$pvalue)){
    return(as.numeric(rep(NA, 4)))
  } else {
    pvals <- x$pvalue
  }
  pvals[x$rep == 0] <- NA
  return(pvals)
}

getphi <- function(x){
  if (is.null(x)) return(numeric(4))
  return(x$statphi$Phi)
}

refactor <- function(strings, factors){
  u_factors <- unique(factors)
  u_factors[match(strings, u_factors)]
}

make_amova_table <- function(am, amt){
  tot <- nrow(am$results)
  res <- data.frame(list(am$results[-tot, c("Df", "Sum Sq")], 
                         Percent = am$componentsofcovariance[-tot, 2],
                         Pval    = rev(amt$pvalue), 
                         Phi     = am$statphi$Phi[-tot]))
  res <- as.matrix(res)
  colnames(res) <- c("d.f.", "Sum of Squares", "Percent variation", "P", 
                     "Phi statistic")
  names(dimnames(res)) <- c("levels", "statistic")
  return(res)
}

make_amova_printable <- function(amtab, amtabcc){
  am_array <- array(dim      = c(dim(amtab), 2),
                    dimnames = c(dimnames(amtab), 
                                 list(c("full", "clone-corrected"))))
  am_array[, , 1] <- amtab
  am_array[, , 2] <- amtabcc
  tabfun <- function(x){
    x <- paste0(paste0(signif(x, 3), collapse = " ("), ")")
    return(x)
  }
  res <- apply(am_array, c(1, 2), tabfun)
  return(res)
}

ci2string <- function(x, sig = 3, colString = "-"){
  if (all(is.na(x))) return("(-)")
  paste0("(", paste(signif(x, sig), collapse = colString),")")
}

add_ci <- function(dtable, ciarray, sig = 3, colString = "-"){
  
  dtable <- data.frame(lapply(dtable, function(x, sig){
    if (is.numeric(x)) x <- signif(x, sig)
    return(x)
  }, sig))
  for (i in colnames(ciarray)){
    to_add <- apply(ciarray[, i, ], 2, ci2string, sig, colString)
    dtable[[i]] <- paste(signif(dtable[[i]], sig), to_add)
  }
  return(dtable)
}

```
Genotypic Diversity
---------

```{r, GD, cache=TRUE}
(GDxYear      <- poppr(ramdat, 
                      sample = 999, 
                      quiet = TRUE, 
                      hist = FALSE))
(GDxRegion    <- poppr(setpop(ramdat, ~ZONE2), 
                      sample = 999, 
                      quiet = TRUE, 
                      hist = FALSE))
(GDxYearccYR  <- poppr(ramdat, 
                      clonecorrect = TRUE, 
                      hier = ~Pop/ZONE2, 
                      sample = 999, 
                      quiet = TRUE, 
                      hist = FALSE))
(GDxRegionccR <- poppr(ramdat, 
                      clonecorrect = TRUE, 
                      hier = ~ZONE2, 
                      sample = 999, 
                      quiet = TRUE, 
                      hist = FALSE))
(Totcc        <- poppr(ramdat, 
                      clonecorrect = TRUE, 
                      hier = NA, 
                      sublist = "Total",
                      sample = 999, 
                      quiet = TRUE, 
                      hist = FALSE))
```


### Confidence intervals for genotypic diversity


```{r, CI, fig.width = 10, fig.height = 10, cache=TRUE}
# By Year:
set.seed(9001)
system.time(year_simp <- boot_ci(ramdat, n = 9999, total = TRUE, 
                      parallel = "multicore", ncpus = 4L))

# By Population:
set.seed(9001)
system.time(zone_simp <- boot_ci(setpop(ramdat, ~ZONE2), n = 9999, total = TRUE, 
                      parallel = "multicore", ncpus = 4L))
```

```{r, add_ci}

write.table(add_ci(GDxYear, year_simp, colString = "-")[-nrow(GDxYear), -length(GDxYear)], 
            file = "GD.csv", 
            na = "-",
            row.names = FALSE,
            sep = ",", col.names = TRUE)
write.table(add_ci(GDxRegion, zone_simp, colString = "-")[-length(GDxYear)],
            file = "GD.csv", 
            na = "-",
            row.names = FALSE,
            sep = ",", col.names = FALSE, append = TRUE)

```


Allelic and Genotypic diversity per locus
--------

```{r, locus_stuff, cache = TRUE, fig.width = 10, fig.height = 10}
(loc_whole <- locus_table(ramdat)) # Whole data set
(loc_geno  <- locus_table(clonecorrect(ramdat, hier = NA))) # By genotype


loc_year_allele <- calc_loc_table(ramdat, ~Pop/ZONE2)
(lya <- plot_loc_table(loc_year_allele) + 
   ggtitle("Allelic Diversity") + labs(list(x = "Year", fill = "Year")))

loc_year_geno <- calc_loc_table(ramdat, ~Pop/ZONE2, lev = "genotype")
(lyg <- plot_loc_table(loc_year_geno) + 
   ggtitle("Genotype Diversity") + labs(list(x = "Year", fill = "Year")))

loc_region_allele <- calc_loc_table(ramdat, ~ZONE2)
(lya <- plot_loc_table(loc_region_allele) + 
   ggtitle("Allelic Diversity") + labs(list(x = "Region", fill = "Region")))

loc_region_geno <- calc_loc_table(ramdat, ~ZONE2, lev = "genotype")
(lyg <- plot_loc_table(loc_region_geno) + 
   ggtitle("Genotype Diversity") + labs(list(x = "Region", fill = "Region")))
```

```{r, create_tables}
loc_region_allele_summary <- melt(loc_region_allele) %>% 
  filter(locus == "mean") %>% 
  mutate(region = ifelse(L1 == "pops", L2, L1)) %>% 
  select(locus, summary, value, region) %>% 
  dcast(region ~ summary) %>%
  arrange(region)
loc_year_allele_summary <- melt(loc_year_allele) %>% 
  filter(locus == "mean") %>% 
  mutate(year = ifelse(L1 == "pops", L2, L1)) %>% 
  select(locus, summary, value, year) %>% 
  dcast(year ~ summary)
```

```{r, print_tables, results = "asis", echo = FALSE}
cat("### Tables\n")
cat("#### Mean Allelic Diversity by Region\n")
knitr::kable(loc_region_allele_summary)

cat("\n#### Mean Allelic Diversity by Year\n")
knitr::kable(loc_year_allele_summary)

cat("\n#### Total Allelic Diversity\n")
knitr::kable(loc_geno)
```



AMOVA
-----

### Including within individual variance

#### Year/Region
```{r, AMOVA, cache=TRUE}
namehierarchy(ramdat) <- ~Year/ZONE2/ZONE1

(year_z2   <- poppr.amova(ramdat, ~Year/ZONE2))
(year_z2cc <- poppr.amova(ramdat, ~Year/ZONE2, clonecorrect = TRUE))

set.seed(9001)
system.time(year_z2.t <- randtest(year_z2, nrepet = 999))
plot(year_z2.t)
set.seed(9001)
system.time(year_z2cc.t <- randtest(year_z2cc, nrepet = 999))
plot(year_z2cc.t)
date()
```

#### Region/Year

```{r, AMOVAzy, cache=TRUE}
namehierarchy(ramdat) <- ~Year/ZONE2/ZONE1

(z2_year   <- poppr.amova(ramdat, ~ZONE2/Year))
(z2_yearcc <- poppr.amova(ramdat, ~ZONE2/Year, clonecorrect = TRUE))

set.seed(9001)
system.time(z2_year.t <- randtest(z2_year, nrepet = 999))
plot(z2_year.t)
set.seed(9001)
system.time(z2_yearcc.t <- randtest(z2_yearcc, nrepet = 999))
plot(z2_yearcc.t)
date()
```

### Excluding within individual variance

#### Year/Region
```{r, AMOVA_WF, cache=TRUE}
namehierarchy(ramdat) <- ~Year/ZONE2/ZONE1

(WF_year_z2   <- poppr.amova(ramdat, ~Year/ZONE2, within = FALSE))
(WF_year_z2cc <- poppr.amova(ramdat, ~Year/ZONE2, clonecorrect = TRUE, 
                          within = FALSE))

set.seed(9001)
system.time(WF_year_z2.t <- randtest(WF_year_z2, nrepet = 9999))
plot(WF_year_z2.t)
set.seed(9001)
system.time(WF_year_z2cc.t <- randtest(WF_year_z2cc, nrepet = 9999))
plot(WF_year_z2cc.t)
date()
```

#### Region/Year
```{r, AMOVAzy_WF, cache=TRUE}
namehierarchy(ramdat) <- ~Year/ZONE2/ZONE1

(WF_z2_year   <- poppr.amova(ramdat, ~ZONE2/Year, within = FALSE))
(WF_z2_yearcc <- poppr.amova(ramdat, ~ZONE2/Year, clonecorrect = TRUE, 
                          within = FALSE))

set.seed(9001)
system.time(WF_z2_year.t <- randtest(WF_z2_year, nrepet = 9999))
plot(WF_z2_year.t)
set.seed(9001)
system.time(WF_z2_yearcc.t <- randtest(WF_z2_yearcc, nrepet = 9999))
plot(WF_z2_yearcc.t)
date()
```

### Writing tables
```{r}
z2year_table   <- make_amova_table(WF_z2_year, WF_z2_year.t)
z2yearcc_table <- make_amova_table(WF_z2_yearcc, WF_z2_yearcc.t)
(z2year_full_table <- make_amova_printable(z2year_table, z2yearcc_table))
write.table(z2year_full_table, file = "zone_by_year.csv", row.names = TRUE,
            col.names = NA, sep = ",")

yearz2_table   <- make_amova_table(WF_year_z2, WF_year_z2.t)
yearz2cc_table <- make_amova_table(WF_year_z2cc, WF_year_z2cc.t)
(yearz2_full_table <- make_amova_printable(yearz2_table, yearz2cc_table))
write.table(yearz2_full_table, file = "year_by_zone.csv", row.names = TRUE,
            col.names = NA, sep = ",")
```


### Without Cape Sebastian or Pistol River South Fork

```{r, noseb, cache = TRUE}
noseb <- popsub(setpop(ramdat, ~ZONE2), blacklist = "HunterCr")

system.time(noseb_year_z2   <- poppr.amova(noseb, ~Year/ZONE2))
system.time(noseb_year_z2cc <- poppr.amova(noseb, ~Year/ZONE2, clonecorrect = TRUE))

set.seed(9001)
system.time(noseb_year_z2.t <- randtest(noseb_year_z2, nrepet = 999))
plot(noseb_year_z2.t)
set.seed(9001)
system.time(noseb_year_z2cc.t <- randtest(noseb_year_z2cc, nrepet = 999))
plot(noseb_year_z2cc.t)
date()

```


```{r, nosebzy, cache = TRUE}
noseb <- popsub(setpop(ramdat, ~ZONE2), blacklist = "HunterCr")

system.time(noseb_z2_year   <- poppr.amova(noseb, ~ZONE2/Year))
system.time(noseb_z2_yearcc <- poppr.amova(noseb, ~ZONE2/Year, clonecorrect = TRUE))

set.seed(9001)
system.time(noseb_z2_year.t <- randtest(noseb_z2_year, nrepet = 999))
plot(noseb_z2_year.t)
set.seed(9001)
system.time(noseb_z2_yearcc.t <- randtest(noseb_z2_yearcc, nrepet = 999))
plot(noseb_z2_yearcc.t)
date()

```


```{r, nosebpr, cache = TRUE}
nosebpr <- selPopSize(noseb, n = 10)

system.time(nosebpr_year_z2   <- poppr.amova(nosebpr, ~Year/ZONE2))
system.time(nosebpr_year_z2cc <- poppr.amova(nosebpr, ~Year/ZONE2, clonecorrect = TRUE))

set.seed(9001)
system.time(nosebpr_year_z2.t <- randtest(nosebpr_year_z2, nrepet = 999))
plot(nosebpr_year_z2.t)
set.seed(9001)
system.time(nosebpr_year_z2cc.t <- randtest(nosebpr_year_z2cc, nrepet = 999))
plot(nosebpr_year_z2cc.t)


```

```{r, nosebprzy, cache = TRUE}
nosebpr <- selPopSize(noseb, n = 10)

system.time(nosebpr_z2_year   <- poppr.amova(nosebpr, ~ZONE2/Year))
system.time(nosebpr_z2_yearcc <- poppr.amova(nosebpr, ~ZONE2/Year, clonecorrect = TRUE))

set.seed(9001)
system.time(nosebpr_z2_year.t <- randtest(nosebpr_z2_year, nrepet = 999))
plot(nosebpr_z2_year.t)
set.seed(9001)
system.time(nosebpr_z2_yearcc.t <- randtest(nosebpr_z2_yearcc, nrepet = 999))
plot(nosebpr_z2_yearcc.t)


```

### Excluding individual variance and Cape Sebastian/PRSF

#### Year/Region
```{r, nosebpr_AMOVA_WF, cache=TRUE}
nosebpr <- resetMLG(nosebpr)

(nosebpr_WF_year_z2   <- poppr.amova(nosebpr, ~Year/ZONE2, within = FALSE))
(nosebpr_WF_year_z2cc <- poppr.amova(nosebpr, ~Year/ZONE2, clonecorrect = TRUE, 
                          within = FALSE))

set.seed(9001)
system.time(nosebpr_WF_year_z2.t <- randtest(nosebpr_WF_year_z2, nrepet = 9999))
plot(nosebpr_WF_year_z2.t)
set.seed(9001)
system.time(nosebpr_WF_year_z2cc.t <- randtest(nosebpr_WF_year_z2cc, nrepet = 9999))
plot(nosebpr_WF_year_z2cc.t)
date()
```

#### Region/Year
```{r, nosebpr_AMOVAzy_WF, cache=TRUE}
namehierarchy(nosebpr) <- ~Year/ZONE2/ZONE1

(nosebpr_WF_z2_year   <- poppr.amova(nosebpr, ~ZONE2/Year, within = FALSE))
(nosebpr_WF_z2_yearcc <- poppr.amova(nosebpr, ~ZONE2/Year, clonecorrect = TRUE, 
                          within = FALSE))

set.seed(9001)
system.time(nosebpr_WF_z2_year.t <- randtest(nosebpr_WF_z2_year, nrepet = 9999))
plot(nosebpr_WF_z2_year.t)
set.seed(9001)
system.time(nosebpr_WF_z2_yearcc.t <- randtest(nosebpr_WF_z2_yearcc, nrepet = 9999))
plot(nosebpr_WF_z2_yearcc.t)
date()
```

```{r}
nosebpr_z2year_table   <- make_amova_table(nosebpr_WF_z2_year, nosebpr_WF_z2_year.t)
nosebpr_z2yearcc_table <- make_amova_table(nosebpr_WF_z2_yearcc, nosebpr_WF_z2_yearcc.t)
(nosebpr_z2year_full_table <- make_amova_printable(nosebpr_z2year_table, nosebpr_z2yearcc_table))
write.table(nosebpr_z2year_full_table, file = "nosebpr_zone_by_year.csv", row.names = TRUE,
            col.names = NA, sep = ",")

nosebpr_yearz2_table   <- make_amova_table(nosebpr_WF_year_z2, nosebpr_WF_year_z2.t)
nosebpr_yearz2cc_table <- make_amova_table(nosebpr_WF_year_z2cc, nosebpr_WF_year_z2cc.t)
(nosebpr_yearz2_full_table <- make_amova_printable(nosebpr_yearz2_table, nosebpr_yearz2cc_table))
write.table(nosebpr_yearz2_full_table, file = "nosebpr_year_by_zone.csv", row.names = TRUE,
            col.names = NA, sep = ",")
```

### Without the coast, cape sebastian, or PRSF
#### Year/Region
```{r, nosebprc_AMOVA_WF, cache=TRUE}
nosebprc <- popsub(setpop(nosebpr, ~ZONE2), blacklist = "Coast")
nosebprc <- resetMLG(nosebprc)

(nosebprc_WF_year_z2   <- poppr.amova(nosebprc, ~Year/ZONE2, within = FALSE))
(nosebprc_WF_year_z2cc <- poppr.amova(nosebprc, ~Year/ZONE2, clonecorrect = TRUE, 
                          within = FALSE))

set.seed(9001)
system.time(nosebprc_WF_year_z2.t <- randtest(nosebprc_WF_year_z2, nrepet = 9999))
plot(nosebprc_WF_year_z2.t)
set.seed(9001)
system.time(nosebprc_WF_year_z2cc.t <- randtest(nosebprc_WF_year_z2cc, nrepet = 9999))
plot(nosebprc_WF_year_z2cc.t)
date()
```

#### Region/Year
```{r, nosebprc_AMOVAzy_WF, cache=TRUE}
namehierarchy(nosebprc) <- ~Year/ZONE2/ZONE1

(nosebprc_WF_z2_year   <- poppr.amova(nosebprc, ~ZONE2/Year, within = FALSE))
(nosebprc_WF_z2_yearcc <- poppr.amova(nosebprc, ~ZONE2/Year, clonecorrect = TRUE, 
                          within = FALSE))

set.seed(9001)
system.time(nosebprc_WF_z2_year.t <- randtest(nosebprc_WF_z2_year, nrepet = 9999))
plot(nosebprc_WF_z2_year.t)
set.seed(9001)
system.time(nosebprc_WF_z2_yearcc.t <- randtest(nosebprc_WF_z2_yearcc, nrepet = 9999))
plot(nosebprc_WF_z2_yearcc.t)
date()
```

```{r}
nosebprc_z2year_table   <- make_amova_table(nosebprc_WF_z2_year, nosebprc_WF_z2_year.t)
nosebprc_z2yearcc_table <- make_amova_table(nosebprc_WF_z2_yearcc, nosebprc_WF_z2_yearcc.t)
(nosebprc_z2year_full_table <- make_amova_printable(nosebprc_z2year_table, nosebprc_z2yearcc_table))
write.table(nosebprc_z2year_full_table, file = "nosebprc_zone_by_year.csv", row.names = TRUE,
            col.names = NA, sep = ",")

nosebprc_yearz2_table   <- make_amova_table(nosebprc_WF_year_z2, nosebprc_WF_year_z2.t)
nosebprc_yearz2cc_table <- make_amova_table(nosebprc_WF_year_z2cc, nosebprc_WF_year_z2cc.t)
(nosebprc_yearz2_full_table <- make_amova_printable(nosebprc_yearz2_table, nosebprc_yearz2cc_table))
write.table(nosebprc_yearz2_full_table, file = "nosebprc_year_by_zone.csv", row.names = TRUE,
            col.names = NA, sep = ",")
```



### Pairwise AMOVA Between Regions

While presenting the AMOVA results, Ebba brought up a very good point that the more recently introduced regions might be biasing the AMOVA results. To assess which regions might have differentiation in a pairwise fashion, a pairwise AMOVA is being run on regions comparing both regions separately per year. 

#### Running the AMOVA by zone
```{r, pair_amova_by_zone, cache = TRUE}
zonePairs    <- pairwise_amova(setpop(ramdat, ~ZONE2), ~ZONE2/Year, quiet = TRUE)
zonePairs.cc <- pairwise_amova(setpop(ramdat, ~ZONE2), ~ZONE2/Year, 
                                quiet = TRUE, clonecorrect = TRUE)
```


```{r, test_amova_by_zone, cache = TRUE}
set.seed(9001) #
system.time(zonePairs.t    <- pairwise_amova_test(zonePairs, nrepet = 999))
set.seed(9001)
system.time(zonePairs.cc.t <- pairwise_amova_test(zonePairs.cc, nrepet = 999))
```

#### Running the AMOVA by year
```{r, pair_amova_by_year, cache = TRUE}
yearPairs    <- pairwise_amova(setpop(ramdat, ~Year), ~Year/ZONE2, quiet = TRUE)
yearPairs.cc <- pairwise_amova(setpop(ramdat, ~Year), ~Year/ZONE2, 
                                quiet = TRUE, clonecorrect = TRUE)
```

```{r, test_amova_by_year, cache = TRUE}
set.seed(9001)
system.time(yearPairs.t    <- pairwise_amova_test(yearPairs, nrepet = 999))
set.seed(9001)
system.time(yearPairs.cc.t <- pairwise_amova_test(yearPairs.cc, nrepet = 999))
```

Pairwise Heatmaps
-------

### By region
```{r, plot_amova_pvals, fig.width = 10, fig.height = 10}
names(zonePairs.t)    <- names(zonePairs)
names(zonePairs.cc.t) <- names(zonePairs.cc)
parray <- array(dim = c(length(zonePairs.t[[1]]$pvalue),
                        length(zonePairs.t),
                        2),
                dimnames = list(Test = zonePairs.t[[1]]$names,
                                Pair = names(zonePairs.t),
                                data = c("full", "clone-censored"))
                )
parray[, , "full"]           <- vapply(zonePairs.t, getpval,
                                       zonePairs.t[[1]]$pvalue)

parray[, , "clone-censored"] <- vapply(zonePairs.cc.t, getpval,
                                       zonePairs.cc.t[[1]]$pvalue)

phiarray <- array(dim = c(length(getphi(zonePairs[[1]])),
                        length(zonePairs),
                        2),
                dimnames = list(Test = rownames(zonePairs[[1]]$statphi),
                                Pair = names(zonePairs),
                                data = c("full", "clone-censored"))
                )
phiarray[, , "full"]           <- vapply(zonePairs, getphi,
                                         getphi(zonePairs[[1]]))

phiarray[, , "clone-censored"] <- vapply(zonePairs.cc, getphi,
                                         getphi(zonePairs.cc[[1]]))



p.df <- tbl_df(melt(parray[3:4, , ])) %>% 
  separate(Pair, c("PairOne", "PairTwo"), sep = " : ")

# Fixing the data so it appaears as an upper-triangle
setpop(ramdat) <- ~ZONE2
p.df$PairOne <- refactor(p.df$PairOne, pop(ramdat))
p.df$PairTwo <- refactor(p.df$PairTwo, pop(ramdat))
# p.df$PairOne <- unique(pop(ramdat))[match(p.df$PairOne, unique(pop(ramdat)))]
# p.df$PairTwo <- unique(pop(ramdat))[match(p.df$PairTwo, unique(pop(ramdat)))]
levels(p.df$Test) <- paste("variations between", c("year within region", "regions"))

# Plotting
ggplot(p.df, aes(x = PairTwo, y = PairOne, fill = -log(value))) + 
  geom_tile() + 
  geom_text(aes(label = value, color = value)) +
  facet_grid(data ~ Test) + 
  scale_fill_continuous(low = "black", high = "orange", name = "p value\n",
                        breaks = c(0.693, 3, 4.6, 6.7),
                        labels = round(exp(-c(0.693, 3, 4.6, 6.8)), 3)
                        ) +
  scale_color_continuous(low = "black", high = "orange", guide = "none") +
  theme_classic() + myTheme + labs(list(x = NULL, y = NULL))

# Creating a matrix to store Phi values
phi <- matrix(nrow = nlevels(p.df$PairOne), ncol = nlevels(p.df$PairOne), 
              dimnames = list(levels(p.df$PairOne), levels(p.df$PairOne)))
p   <- phi
phi[lower.tri(phi)] <- phiarray[4, , 2] # Clone corrected (upper triangle)
phi <- t(phi) # Transposition to final form
phi[lower.tri(phi)] <- phiarray[4, , 1] # Raw data (lower triangle)
phi
p[lower.tri(p)]     <- parray[4, , 2]
p <- t(p)
p[lower.tri(p)]     <- parray[4, , 1]
p
write.table(phi, file = "phi_by_zone.csv", sep = ",", row.names = TRUE,
            col.names = NA)
```


### By year


```{r, plot_amova_pvals_year, fig.width = 15, fig.height = 15}
names(yearPairs.t)    <- names(yearPairs)
names(yearPairs.cc.t) <- names(yearPairs.cc)
parray_year <- array(dim = c(length(yearPairs.t[[1]]$pvalue),
                        length(yearPairs.t),
                        2),
                dimnames = list(Test = yearPairs.t[[1]]$names,
                                Pair = names(yearPairs.t),
                                data = c("full", "clone-censored"))
                )

parray_year[, , "full"]           <- vapply(yearPairs.t, getpval, 
                                            yearPairs.t[[1]]$pvalue)

parray_year[, , "clone-censored"] <- vapply(yearPairs.cc.t, getpval,
                                            yearPairs.cc.t[[1]]$pvalue)

phiarray_year <- array(dim = c(length(getphi(yearPairs[[1]])),
                        length(yearPairs),
                        2),
                dimnames = list(Test = rownames(yearPairs[[1]]$statphi),
                                Pair = names(yearPairs),
                                data = c("full", "clone-censored"))
                )
phiarray_year[, , "full"]           <- vapply(yearPairs, getphi,
                                         getphi(yearPairs[[1]]))

phiarray_year[, , "clone-censored"] <- vapply(yearPairs.cc, getphi,
                                         getphi(yearPairs.cc[[1]]))

py.df <- tbl_df(melt(parray_year[3:4, , ])) %>% 
  separate(Pair, c("PairOne", "PairTwo"), sep = " : ")

# Fixing the data so it appaears as an upper-triangle
setpop(ramdat) <- ~Year
py.df$PairOne <- refactor(py.df$PairOne, pop(ramdat))
py.df$PairTwo <- refactor(py.df$PairTwo, pop(ramdat))
# py.df$PairOne <- unique(pop(ramdat))[match(py.df$PairOne, unique(pop(ramdat)))]
# py.df$PairTwo <- unique(pop(ramdat))[match(py.df$PairTwo, unique(pop(ramdat)))]
levels(py.df$Test) <- paste("variations between", c("regions within years", "years"))

# Plotting
ggplot(py.df, aes(x = PairTwo, y = PairOne, fill = -log(value))) + 
  geom_tile() + 
  geom_text(aes(label = value, color = value)) +
  facet_grid(data ~ Test) + 
  scale_fill_continuous(low = "black", high = "orange", name = "p value\n",
                        breaks = c(0.693, 3, 4.6, 6.7),
                        labels = round(exp(-c(0.693, 3, 4.6, 6.8)), 3)
                        ) +
  scale_color_continuous(low = "black", high = "orange", guide = "none") +
  theme_classic() + myTheme + labs(list(x = NULL, y = NULL))

phi_year <- matrix(nrow = nlevels(py.df$PairOne), ncol = nlevels(py.df$PairOne), 
              dimnames = list(levels(py.df$PairOne), levels(py.df$PairOne)))

phi_year[lower.tri(phi_year)] <- phiarray_year[4, , 2]
phi_year <- t(phi_year)
phi_year[lower.tri(phi_year)] <- phiarray_year[4, , 1]
phi_year
write.table(phi_year, file = "phi_by_year.csv", sep = ",", row.names = TRUE,
            col.names = NA)
```

