---
title: "Population structure"
author: "Zhian N. Kamvar"
date: "September 30, 2014"
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
    toc_depth: 2
---

Purpose
-------

This document will explore population clustering with bootstrapping. This will
demonstrate to some degree the differentiation between populations.

Required packages and data
--------

```{r}
library(PramCurry)
library(reshape2)
library(ggplot2)
library(ape)
library(poppr)
library(adegenet)
library(igraph)
options(stringsAsFactors = FALSE)
data(ramdat)
data(pop_data)
data(myPal)
sessionInfo()
```


Custom Function
------
```{r}
neiboot <- function(x, sample = 1000, color_by = "black"){
  x <- genind2genpop(x, quiet = TRUE)
  outTree <- aboot(x, sample = sample, cutoff = 50, quiet = TRUE, tree = "nj",
                   showtree = FALSE)
  plot.phylo(outTree, xpd = TRUE, font = 2, cex = 0.8, type = "u",
             label.offset = 0.005, adj = 0, tip.col = color_by)
  nodelabels(outTree$node.label, xpd = TRUE, frame = "n", cex = 0.8, font = 3)#,
#              adj = c(1.3, -0.5))
#   axisPhylo(3)
  add.scale.bar(lwd = 5)
  return(outTree)
}
```


Population level dendrograms
--------

First, I will take a look at what the different population factors reveal.

```{r, pop, fig.height = 15, fig.width = 10, cache=TRUE}
par(mfrow = c(3, 1))
system.time(neiboot(ramdat))
system.time(neiboot(setpop(ramdat, ~ZONE2)))
system.time(neiboot(setpop(ramdat, ~ZONE1)))
par(mfrow = c(1, 1))
```

We can see that Cape Sebastian/Hunter's Creek is creating strong outgroups.
I will run the analyses on the different zones without these populations.

```{r, pop_no_seb, fig.height = 10, fig.width = 10, cache = TRUE}
par(mfrow = c(2, 1))
system.time(neiboot(popsub(setpop(ramdat, ~ZONE2), blacklist = "HunterCr")))
system.time(neiboot(popsub(setpop(ramdat, ~ZONE1), blacklist = "CapeSeb")))
par(mfrow = c(1, 1))
```

### Population dendrograms by year

Now we want to analyze the populations by year. These will be colored by Zone2
since that is the top level of the hierarchy.

#### With Cape Sebastian

```{r, popyear, fig.height = 15, fig.width = 10, cache = TRUE}
par(mfrow = c(2, 1))

Z2    <- setpop(ramdat, ~ZONE2)
Z2Pal <- char2pal(Z2@pop.names)
setpop(ramdat) <- ~ZONE2/Pop
set.seed(9001)
system.time(outtree <- neiboot(ramdat, sample = 10000,
                               color_by = Z2Pal[sub("_.+?$", "", ramdat@pop.names)]))

setpop(ramdat) <- ~ZONE2/ZONE1/Pop
set.seed(9001)
system.time(neiboot(ramdat, sample = 10000, 
                    color_by = Z2Pal[sub("_.+?$", "", ramdat@pop.names)]))

par(mfrow = c(1, 1))

write.nexus(outtree, file = "zone_by_year.nex")
```

#### Without Cape Sebastian

```{r, popyear_no_seb, fig.height = 15, fig.width = 10, cache = TRUE}

par(mfrow = c(2, 1))

ramdat <- popsub(setpop(ramdat, ~ZONE2/Pop), blacklist = "HunterCr_2011")
Z2     <- setpop(ramdat, ~ZONE2)
Z2Pal  <- char2pal(Z2@pop.names)
setpop(ramdat) <- ~ZONE2/Pop
system.time(neiboot(ramdat, color_by = Z2Pal[sub("_.+?$", "", ramdat@pop.names)]))

ramdat <- popsub(setpop(ramdat, ~ZONE1/Pop), blacklist = "CapeSeb_2011")
setpop(ramdat) <- ~ZONE2/ZONE1/Pop
system.time(neiboot(ramdat, color_by = Z2Pal[sub("_.+?$", "", ramdat@pop.names)]))

par(mfrow = c(1, 1))


```